A full approximation scheme multilevel method \\ for nonlinear variational inequalities

### 2023-07-24 paper/fascd.tex
Minor edits

Here we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach.  Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method extends the \cite{GraeserKornhuber2009,Tai2003} algorithms in the following directions:
\pef{These aren't all really extensions. Maybe we could say `enjoys the following advantages'? We distinguish it from prior work below, so this section does not need to do that.}
\renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
\begin{enumerate}
\item The FAS coarse corrections are adapted for general VI problems. The needed computations require neither an objective function nor linearity of the residual operator.
\item Box constraints (i.e.~both upper and lower obstacles) are allowed.
\item As with the original CD method \cite{Tai2003}, but unlike some implementations \cite[for example]{GraeserKornhuber2009}, and semismooth methods generally \cite{BensonMunson2006,Ulbrich2011}, the method can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
\pef{I confess I find this explicit calling out of other algorithms a bit uncomfortable/too much. I wonder if the following is better?}
\item The method is strictly admissible, and so can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
\item V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}) are implemented.
\item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
\item When used in conjunction with a reduced-space active set method as a smoother, we demonstrate that F-cycle FASCD is an optimal solver for the different VI examples in Section \ref{sec:results}.  The work of the smoother is proportional to the number of unknowns on a given level because the number of Newton and Krylov iterations is fixed.
\end{enumerate}

@@

The observation that up-smoothing is more efficient in a multilevel CD context, which arises because of constraint set distinctions, seems to be new, but compare comments on $V(1,1)$ cycles in \cite{GraeserKornhuber2009,Tai2003}.  The increased efficiency is particularly shown in Example \ref{ex:results:plap}.  In summary, the distinction is that construction of constraint sets via monotone injection generates up-smoothing sets which can telescope the down-smoothing sets (Lemma \ref{lem:upwardadmissibility}); we call this an \emph{incomplete CD}.  This structure can be exploited by a new CD iteration, algorithm \pr{icd-tele} in Section \ref{sec:cdmultilevel}.

@@

\subsection*{Acknowledgements} \label{sec:acknowledgements}  EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks. PEF was supported by the Engineering and Physical Sciences Research Council [EPSRC grants EP/R029423/1 and EP/W026163/1]. This work used the ARCHER2 UK National Supercomputing Service (https://www.archer2.ac.uk). PEF thanks Jack D.~Betteridge for assistance with the numerical experiments of Section \ref{ex:results:sia}.

>>>>

Here we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach.  Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method extends the \cite{GraeserKornhuber2009,Tai2003} algorithms in the following directions:
\pef{These aren't all really extensions. Maybe we could say `enjoys the following advantages'? We distinguish it from prior work below, so this section does not need to do that.}
\renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
\begin{enumerate}
\item The FAS coarse corrections are adapted for general VI problems. The needed computations require neither an objective function nor linearity of the residual operator.
\item Box constraints (i.e.~both upper and lower obstacles) are allowed.
\item As with the original CD method \cite{Tai2003}, but unlike some implementations \cite[for example]{GraeserKornhuber2009}, and semismooth methods generally \cite{BensonMunson2006,Ulbrich2011}, the method can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
\pef{I confess I find this explicit calling out of other algorithms a bit uncomfortable/too much. I wonder if the following is better?}
\item The method is strictly admissible, and so can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
\item V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}) are proposed.
\item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
\item When used in conjunction with a reduced-space active set method as a smoother, we demonstrate that F-cycle FASCD is an optimal solver for the different VI examples in Section \ref{sec:results}.  The work of the smoother is proportional to the number of unknowns on a given level because the number of Newton and Krylov iterations is fixed.
\end{enumerate}

@@

The observation that up-smoothing is more efficient in a multilevel CD context, which arises because of constraint set distinctions, seems to be new, but compare comments on $V(1,1)$ cycles in \cite{GraeserKornhuber2009,Tai2003}.  The increased efficiency is particularly shown in Example \ref{ex:results:plap}.  In summary, the distinction is that construction of constraint sets via monotone injection generates up-smoothing sets which can telescope the down-smoothing sets (Lemma \ref{lem:upwardadmissibility}); we call this an \emph{incomplete CD}.  This structure can be exploited by a new CD iteration, algorithm \pr{icd-tele} in Section \ref{sec:cdmultilevel}.
\pef{This last paragraph feels like a flat way to end. Perhaps we could highlight the importance of ICD-TELE in the text, and remove this last paragraph?}

@@

\subsection*{Acknowledgements} \label{sec:acknowledgements}  EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks. PEF was supported by the Engineering and Physical Sciences Research Council [EPSRC grants EP/R029423/1 and EP/W026163/1]. This work used the ARCHER2 UK National Supercomputing Service (https://www.archer2.ac.uk). PEF thanks Jack D.~Betteridge for assistance with the numerical experiments of Example \ref{ex:results:sia}.
